---
title: "Aquila Analysis"
author: "Luca Torello Pianale"
date: "March 2023"
header-includes:
  - \usepackage{fvextra}
  - \DefineVerbatimEnvironment{Highlighting}{Verbatim}{breaksymbolleft={},showspaces = false,showtabs = false,breaklines,commandchars=\\\{\}}
output:
  pdf_document: default
  html_document: default
---

### SCOPE OF THE MARKDOWN

This code is used to analyse the data coming from Aquila runs.


```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, 
                      tidy = "styler", fig.width = 10, fig.height = 12)

```
\newpage  

# LIBRARIES

Automated installation (if needed) and loading of libraries required for this script.  
The packages "tidyverse" and "rstatix" are needed for smooth data and statistical analysis. 
The packages "readxl" and "writexl" are needed to import and export the data.  
The packages "ggpubr" and "ggplot2" are needed for plotting. 
The packages "deSolve" and "growthrates" are needed for specific growth rates and lag phases.  

```{r Libraries}

#R Version used
R.Version()[["version.string"]]

#Libraries
requiredPackages <- c("tidyverse", "ggplot2", "ggpubr", "readxl", "writexl", "rstatix", "deSolve", "growthrates")

ipak <- function(pkg){
        new.pkg <- pkg[!(pkg %in% installed.packages()[, "Package"])]
        if (length(new.pkg))
                install.packages(new.pkg, dependencies = TRUE)
        sapply(pkg, require, character.only = TRUE)
}

ipak(requiredPackages)
```
\newpage  

# DATA ANALYSIS

## Define general info

Assign the file path in the computer to the variables "folder" and "file".  
The R file should be in the same folder to use command getwd() in this context! Moreover, the files should be in .xlsx format. Note that, sometimes, files will not load in R if it is open in excel!

FYI: Soft Wood (Spruce), Hard Wood (Birch), Non-Woody (oat & corn - annual, bagasse & wheat - perennial)

```{r Data_preparation}

#Paths
folder <- getwd()
setwd(folder)
input_folder <- paste0(folder, "/Input_data")
input_files <- paste0("/", list.files(input_folder)[endsWith(list.files(input_folder), "_OK.xlsx")])

#General info
info <- list(media = c("Delft", "SLRH50", "HGSH50", "BiH50",
                       "WSH50", "SBH50", "CSH50", "OHH50"),
             colors_media = c("#5A5A5A", "#c2a5cf", "#A45BBD", "#7B5FCA",
                              "#e5f5f9", "#99d8c9", "#41ae76", "#006d2c"),
             materials = c("All", "NON-Woody", "Woody"),
             strains = c("CEN.PK113-7D", "EthanolRed", "PE2"),
             colors_strains = c("#333333", "#FC8D62", "#89ADFF"),
             Functions = c("mumax", "lag", "yCDW", "yEth", "yGly", "yAAc"),
             Functions_units_x =  c(lag = "Lag Phase (h)",
                                   mumax = "Specific Growth\nRate (1/h)",
                                   yEth = "Ethanol Yield\n(g/g)",
                                   yGly = "Glycerol Yield\n(g/g)",
                                   yAAc = "Acetic Acid Yield\n(g/g)",
                                   yCDW = "Cell Mass Yield\n(g/g)"),
             Functions_units_y =  c(lag = "Lag Phase (h)",
                                   mumax = "Specific Growth\nRate (1/h)",
                                   yEth = "Ethanol Yield\n(g/g)",
                                   yGly = "Glycerol Yield\n(g/g)",
                                   yAAc = "Acetic Acid Yield\n(g/g)",
                                   yCDW = "Cell Mass Yield\n(g/g)"))
```

## Load Data

```{r Data_preparation}

#Load summary
summary <- read_excel(paste0(input_folder, "/Layout.xlsx"), 
                      sheet = "layout", col_names = TRUE)

#Load data
all_data <- data.frame()

for (k in input_files) {
  list <- excel_sheets(paste0(input_folder, k))
  df2 <- data.frame()
  for (i in list) {
    df1 <- read_excel(paste0(input_folder, k), sheet = i, col_names = TRUE) %>%
      as.data.frame() %>%
      mutate(strain = case_when(startsWith(k, "/CEN") ~ "CEN.PK113-7D",
                                startsWith(k, "/Eth") ~ "EthanolRed",
                                startsWith(k, "/PE") ~ "PE2")) %>%
      cbind(position = gsub("(.+)-(.+)", "\\2", i),
            station = gsub("(.+)-(.+)", "\\1", i)) 
    
    df2 <- bind_rows(df2, df1)
    rm(df1)
  }
  
  df2 <- merge(df2, summary, all.x = T) %>%
    select(-c(position, station))
  
  all_data <- bind_rows(all_data, df2) 
  rm(df2, list)
  }

#Reorganise data
all_data <- all_data %>%
  filter(medium %in% info[["media"]]) %>%
  rename(value = backscatter) %>%
  group_by(strain, medium, replicate) %>%
  arrange(time, .by_group = T) %>%
  mutate(value = value/first(value), #Normalise by first value
         across(where(is.numeric), ~ round(.,4)))

#Clean environment
rm(summary, i, k, input_files)
```
\newpage  

# GROWTH FunctionS

## Mu and Lag

Computing specific growth rates (mumax) and lag phases + statistical analysis.

```{r Growth_Check}

#Analysis of the growth curves for mumax. Note: in the output graphs the inflection point and the fitting curve are shown.
many_spline_fits <- all_splines(value ~ time | strain + medium + replicate,
                                data = subset(all_data, time > 4), spar = 0.3) 
#par(mfrow = c(8, 8)) 
#par(mar = rep(1, 4)) 
#plot(many_spline_fits) 
growth_param <- results(many_spline_fits) %>%
  remove_rownames       #Saving the results


#Calculate lambda (lag phase): the coordinates of the inflection point have been extracted and used to calculate the tangent and subsequently the x value corresponding to the lag phase.

inf_points <- NULL

for (i in 1:length(many_spline_fits@fits)){
  tmp <- many_spline_fits@fits[[i]]@xy
  inf_points <- rbind(inf_points, tmp)
  rm(tmp)
}

inf_points <- as.data.frame(inf_points)

growth_param <- bind_cols(growth_param, inf_points) %>%
  rename(., x = V1, y = V2) %>%
  mutate(lag = ((log10(y0) - log10(y)) / mumax) + x) %>%
  select(-c(x, y, y0, r2)) %>%
  pivot_longer(c(lag, mumax), names_to = "Function", values_to = "value")

#Rearrange data 
growth_param$value[growth_param$strain == "CEN.PK113-7D" & growth_param$medium %in% c("HGSH50", "BiH50", "SLRH50") & growth_param$Function == "lag" & growth_param$replicate == 1] <- 48
growth_param$value[growth_param$strain == "CEN.PK113-7D" & growth_param$medium %in% c("HGSH50", "BiH50", "SLRH50") & growth_param$Function == "lag" & growth_param$replicate == 2] <- 47.9
growth_param$value[growth_param$strain == "CEN.PK113-7D" & growth_param$medium %in% c("HGSH50", "BiH50", "SLRH50") & growth_param$Function == "lag" & growth_param$replicate == 3] <- 47.99

growth_param$value[growth_param$strain == "CEN.PK113-7D" & growth_param$medium %in% c("HGSH50", "BiH50", "SLRH50") & growth_param$Function == "mumax" & growth_param$replicate == 1] <- 0.001
growth_param$value[growth_param$strain == "CEN.PK113-7D" & growth_param$medium %in% c("HGSH50", "BiH50", "SLRH50") & growth_param$Function == "mumax" & growth_param$replicate == 2] <- 0.0011
growth_param$value[growth_param$strain == "CEN.PK113-7D" & growth_param$medium %in% c("HGSH50", "BiH50", "SLRH50") & growth_param$Function == "mumax" & growth_param$replicate == 3] <- 0.0012


growth_param$value[growth_param$strain == "PE2" & growth_param$medium %in% c("OHH50", "WSH50", "BiH50", "SLRH50") & growth_param$Function == "lag" & growth_param$replicate == 1] <- 48
growth_param$value[growth_param$strain == "PE2" & growth_param$medium %in% c("OHH50", "WSH50", "BiH50", "SLRH50") & growth_param$Function == "lag" & growth_param$replicate == 2] <- 47.9
growth_param$value[growth_param$strain == "PE2" & growth_param$medium %in% c("OHH50", "WSH50", "BiH50", "SLRH50") & growth_param$Function == "lag" & growth_param$replicate == 3] <- 47.99

growth_param$value[growth_param$strain == "PE2" & growth_param$medium %in% c("OHH50", "WSH50", "BiH50", "SLRH50") & growth_param$Function == "mumax" & growth_param$replicate == 1] <- 0.001
growth_param$value[growth_param$strain == "PE2" & growth_param$medium %in% c("OHH50", "WSH50", "BiH50", "SLRH50") & growth_param$Function == "mumax" & growth_param$replicate == 2] <- 0.0011
growth_param$value[growth_param$strain == "PE2" & growth_param$medium %in% c("OHH50", "WSH50", "BiH50", "SLRH50") & growth_param$Function == "mumax" & growth_param$replicate == 3] <- 0.0012


growth_param$value[growth_param$strain == "EthanolRed" & growth_param$medium %in% c("SLRH50") & growth_param$Function == "lag" & growth_param$replicate == 1] <- 48
growth_param$value[growth_param$strain == "EthanolRed" & growth_param$medium %in% c("SLRH50") & growth_param$Function == "lag" & growth_param$replicate == 2] <- 47.9
growth_param$value[growth_param$strain == "EthanolRed" & growth_param$medium %in% c("SLRH50") & growth_param$Function == "lag" & growth_param$replicate == 3] <- 47.99

growth_param$value[growth_param$strain == "EthanolRed" & growth_param$medium %in% c("SLRH50") & growth_param$Function == "mumax" & growth_param$replicate == 1] <- 0.001
growth_param$value[growth_param$strain == "EthanolRed" & growth_param$medium %in% c("SLRH50") & growth_param$Function == "mumax" & growth_param$replicate == 2] <- 0.0011
growth_param$value[growth_param$strain == "EthanolRed" & growth_param$medium %in% c("SLRH50") & growth_param$Function == "mumax" & growth_param$replicate == 3] <- 0.0012

#Statistical comparison of different media with respect to the control "Delft".
Stats_medium <- compare_means(value~medium, data = growth_param, 
                              group.by = c("strain", "Function"), 
                              method = "t.test", ref.group = "Delft") %>% 
  add_significance("p.adj") 

#Clean environment
rm(i, inf_points, many_spline_fits)
```

## Yields

Computing yields.

```{r Yields}

#Cell mass
cell_mass <- read_excel(paste0(input_folder, "/Data_HPLC+CDW.xlsx"), 
                        sheet = "cdw", col_names = TRUE) %>%
  select(strain, timepoint, medium, replicate, cellmass_gL)

df1 <- subset(cell_mass, timepoint == "start") %>%
  group_by(strain) %>%
  transmute(CDW_i = mean(cellmass_gL)) %>%
  distinct()

df2 <- subset(cell_mass, timepoint == "end") %>%
  rename(CDW_f = cellmass_gL) %>%
  select(-c(timepoint))

cell_mass <- merge(df1, df2)


#Media components
yields <- read_excel(paste0(input_folder, "/Data_HPLC+CDW.xlsx"), 
                     sheet = "media_composition", col_names = TRUE)

df1 <- subset(yields, timepoint == "start") %>%
  select(-timepoint) %>%
  rename_with(~paste0(.,"_i"), -c(strain, medium, replicate)) %>%
  select(-replicate)

df2 <- subset(yields, timepoint == "end") %>%
  select(-timepoint) %>%
  rename_with(~paste0(.,"_f"), -c(strain, medium, replicate)) 

yields <- merge(df2, df1, all.x = T) %>%
  merge(., cell_mass)

yields[is.na(yields)] <- 0

#Compute yields
yields <- yields %>%
  mutate(hexoses = Glu_i-Glu_f+Gal_i-Gal_f+Man_i-Man_f) %>%
  transmute(strain = strain, 
            medium = medium, 
            replicate = replicate, 
            yEth = (Eth_f-Eth_i)/hexoses,
            yGly = (Gly_f-Gly_i)/hexoses,
            yCDW = (CDW_f-CDW_i)/hexoses) %>%
  pivot_longer(., cols = names(.)[startsWith(names(.), "y")], 
               names_to = "Function", values_to = "value") 


#Merge with mu and lag
growth_param <- bind_rows(growth_param, yields) %>%
  mutate(material = case_when(medium %in% c("SLRH50", "HGSH50", "BiH50") ~ "Woody",
                              medium %in% c("CSH50", "OHH50", "WSH50", "SBH50") ~ "NON-Woody")) %>%
  rbind(., mutate(., material = "All")) %>%
  drop_na(., material)

#Clean environment
rm(cell_mass, yields, list = ls()[startsWith(ls(), "df")])
```
\newpage  

# ROBUSTNESS

```{r Robustness}

#Compute Robustness for all the functions
R_growth <- growth_param %>%
  ungroup() %>%
  group_by(strain, Function, material) %>%
  transmute(mean_value = mean(value), 
            sd_value = sd(value),
            R = -sd_value^2/mean_value,
            value = value) %>%
  ungroup() %>%
  group_by(Function, material) %>%
  transmute(strain = strain, 
            mean_value = mean_value, 
            sd_value = sd_value,
            R = R,
            mean_strains = mean(value),
            R_norm = R/mean_strains)


R_s <- growth_param %>%
  ungroup() %>%
  group_by(medium, Function, material) %>%
  transmute(mean_value = mean(value), 
            sd_value = sd(value),
            R = -sd_value^2/mean_value,
            value = value) %>%
  ungroup() %>%
  group_by(Function, material) %>%
  transmute(medium = medium, 
            mean_value = mean_value, 
            sd_value = sd_value,
            R = R,
            mean_media = mean(value),
            R_norm = R/mean_media)
```

# EXPORT DATA.

AA.  

```{r Export_data}

#Export robustness data frames in an .xlsx file (same file, different sheets)
name <- "/Data_Analysed"
sheets <- list("Data" = all_data, 
               "Growth_Param" = growth_param, 
               "Robustness" = R_growth) 
write_xlsx(sheets, path = paste0(folder, name, ".xlsx"), use_zip64 = T)

#Clean environment
rm(sheets, name)
```
\newpage
# PLOTTING.  

## Data Re-organisation  
<br>
```{r Data_reorganisation, echo = FALSE}

#Re-organising data frames for plotting
for (i in names(Filter(is.data.frame, as.list(.GlobalEnv)))) {
  assign(i, eval(as.symbol(i)) %>%
           mutate(across(any_of(c("strain")), 
                         ~factor(strain, levels = info[["strains"]])),
                  across(any_of(c("material")), 
                         ~factor(material, levels = info[["materials"]])),
                  across(any_of(c("medium")), 
                         ~factor(medium, levels = info[["media"]])),
                  across(any_of(c("Function")), 
                         ~factor(Function, levels = info[["Functions"]]))))
  }
```
\newpage  
## Growth curves

```{r Growth_curves}

lineplots <- ggplot(data = all_data %>% 
                      mutate(value = log(value), 
                             medium = factor(medium, levels = info[["media"]]))) + 
  aes(x = time, y = value, colour = strain) +
  stat_summary(fun = mean, geom = "line", linewidth = 1) +
  stat_summary(fun.data = mean_sdl, geom = "errorbar", 
               fun.args = list(mult = 1), width = 0.05, alpha = 0.1) +
  scale_x_continuous(breaks = seq(from = 0, 
                                  to = max(unique(all_data$time)), 
                                  by = 12)) +
  scale_y_continuous(labels = function(x) format(x, nsmall = 1)) +
  scale_color_manual(breaks = info[["strains"]], 
                     values = info[["colors_strains"]]) +
  labs(y = "ln(Norm. Scattered Light) (a.u.)\n", x = "\nTime (h)") +
  facet_wrap(~medium, nrow = 2) +
  theme_light()+ 
  theme(legend.position = "top",
        axis.title = element_text(size = 11),
        axis.text = element_text(size = 11),
        panel.grid.minor = element_blank(),
        panel.spacing.x = unit(1, "lines"), 
        strip.background = element_rect(colour = "white", fill = "black"),
        strip.text = element_text(size = 11, face = "bold.italic"),
        plot.margin = unit(c(0.6, 0.6, 0.6, 0.6), "cm"))

#lineplots
```
\newpage  
## Barplots for functions

```{r Functions}

barplots <- ggplot(data = subset(growth_param, material == "All")) + 
  aes(x = medium, y = value) +
  stat_summary(aes(fill = medium), fun = mean, geom = "bar", 
               position = position_dodge(width = 0.9),
              # color = "black", linewidth = 0.5
               ) +
  stat_summary(fun.data = mean_sdl, geom = "errorbar", 
               fun.args = list(mult = 1),
               width = 0.1, position = position_dodge(width = 0.9)) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.1)), 
                     labels = function(x) format(x, nsmall = 1),
                     breaks = scales::extended_breaks(n = 3)) +
  scale_fill_manual(breaks = info[["media"]], 
                    values = info[["colors_media"]]) +
  facet_grid(cols = vars(strain),
             rows = vars(Function),
             switch = "y", scales = "free",
             labeller = labeller(Function = info[["Functions_units_y"]])) +
  labs(x = "", y = "") +
  guides(fill = guide_legend(ncol = 1)) +
  theme_light() +
  theme(legend.position = "right",
        axis.text.x = element_blank(),
        axis.text.y = element_text(size = 9),
        axis.ticks.x = element_blank(),
        strip.placement.y = "outside",
        strip.background.x = element_rect(colour = "black", fill = "black"),
        strip.background.y = element_blank(),
        strip.text.x = element_text(size = 9, face = "bold.italic", colour = "white"),
        strip.text.y = element_text(size = 9, colour = "black"),
        panel.spacing.x = unit(1, "lines"),
        panel.spacing.y = unit(0.7, "lines"), 
        panel.grid.minor = element_blank(), 
        panel.grid.major.x = element_blank(),
        plot.margin = unit(rep(0.3, 4), "cm"))

#barplots
```
\newpage  
## Robustness

```{r Robustness}

Rc_Perf <- ggplot(data = subset(R_growth, material == "All")) +
  geom_point(aes(x = mean_value, y = R_norm, color = strain), size = 3) +
  geom_errorbarh(aes(xmax = mean_value + sd_value, 
                     xmin = mean_value - sd_value, 
                     y = R_norm, color = strain), height = 0) +
  labs(x = NULL, 
       y = "R(c) (a.u.)\n") +
  scale_y_continuous(expand = expansion(mult = c(0.1, 0.1)), 
                     labels = function(x) format(x, nsmall = 1),
                     limits = c(NA, 0),
                     breaks = scales::extended_breaks(n = 3)) +
  scale_x_continuous(breaks = scales::extended_breaks(n = 4, 
                                                      w = c(0.25, 0.2, 0.9, 0.05))) +
  facet_wrap(~Function, ncol = 3,
             scales = "free",  strip.position = "bottom",
             labeller = labeller(Function = info[["Functions_units_x"]])) +
  scale_color_manual(breaks = info[["strains"]], 
                     values = info[["colors_strains"]]) +
  theme_light() +
  theme(legend.position = c(0.87, 0.18), #(1,1) top right, (0,0) bottom left
        legend.background = element_rect(linewidth = 0.2, 
                                         colour = "black",
                                         linetype = "solid"),
        axis.text.x = element_text(size = 9),
        axis.text.y = element_text(size = 9),
        strip.placement = "outside",
        strip.background.y = element_rect(colour = "black", fill = "black"),
        strip.background.x = element_blank(),
        strip.text.y = element_text(size = 9, face = "bold.italic", colour = "white"),
        strip.text.x = element_text(size = 9, colour = "black"),
        panel.spacing.x = unit(1, "lines"),
        panel.spacing.y = unit(1.5, "lines"), 
        panel.grid.minor = element_blank(), 
        plot.margin = unit(rep(0.3, 4), "cm"))

Rs_Perf <- ggplot(data = subset(R_s, material == "All")) +
  geom_point(aes(x = mean_value, y = R_norm, color = medium), size = 3) +
  geom_errorbarh(aes(xmax = mean_value + sd_value, 
                     xmin = mean_value - sd_value, 
                     y = R_norm, color = medium), height = 0) +
  labs(x = NULL, 
       y = "R(s) (a.u.)\n") +
  scale_y_continuous(expand = expansion(mult = c(0.1, 0.1)), 
                     labels = function(x) format(x, nsmall = 1),
                     limits = c(NA, 0),
                     breaks = scales::extended_breaks(n = 3)) +
  scale_x_continuous(breaks = scales::extended_breaks(n = 4, 
                                                      w = c(0.25, 0.2, 0.9, 0.05))) +
  facet_wrap(~Function, ncol = 3,
             scales = "free",  strip.position = "bottom",
             labeller = labeller(Function = info[["Functions_units_x"]])) +
  scale_color_manual(breaks = info[["media"]], 
                     values = info[["colors_media"]]) +
  guides(color = guide_legend(ncol = 2)) + 
  theme_light() +
  theme(legend.position = c(0.85, 0.18), #(1,1) top right, (0,0) bottom left
        legend.background = element_rect(linewidth = 0.2, 
                                         colour = "black",
                                         linetype = "solid"),
        axis.text.x = element_text(size = 9),
        axis.text.y = element_text(size = 9),
        strip.placement = "outside",
        strip.background.y = element_rect(colour = "black", fill = "black"),
        strip.background.x = element_blank(),
        strip.text.y = element_text(size = 9, face = "bold.italic", colour = "white"),
        strip.text.x = element_text(size = 9, colour = "black"),
        panel.spacing.x = unit(1, "lines"),
        panel.spacing.y = unit(1.5, "lines"), 
        panel.grid.minor = element_blank(), 
        plot.margin = unit(rep(0.3, 4), "cm"))

Rc_Perf_material <- ggplot(data = subset(R_growth, material != "All")) +
  geom_point(aes(x = mean_value, y = R_norm, color = strain, shape = ), size = 3) +
  geom_errorbarh(aes(xmax = mean_value + sd_value, 
                     xmin = mean_value - sd_value, 
                     y = R_norm, color = strain), height = 0) +
  labs(x = NULL, y = "R(c) (a.u.)\n") +
  scale_y_continuous(expand = expansion(mult = c(0.1, 0.1)), 
                     labels = function(x) format(x, nsmall = 1),
                     limits = c(NA, 0)) +
  facet_grid(cols = vars(Function), rows = vars(material), 
             scales = "free",  switch = "x",
             labeller = labeller(Function = info[["Functions_units_x"]])) +
  scale_color_manual(breaks = info[["strains"]], 
                     values = info[["colors_strains"]]) +
  theme_light() +
  theme(legend.position = "top", 
        axis.text.x = element_text(size = 9, angle = 45, hjust = 1),
        axis.text.y = element_text(size = 9),
        strip.placement = "outside",
        strip.background.y = element_rect(colour = "black", fill = "black"),
        strip.background.x = element_blank(),
        strip.text.y = element_text(size = 9, face = "bold.italic", colour = "white"),
        strip.text.x = element_text(size = 9, colour = "black"),
        panel.spacing.x = unit(1.3, "lines"),
        panel.spacing.y = unit(0.7, "lines"), 
        panel.grid.minor = element_blank(), 
        plot.margin = unit(rep(0.3, 4), "cm"))

```
\newpage  
## Saving Figures

```{r Save_Figures}

#Figures
ggsave(filename = "Rc_Perf.png", 
       plot = Rc_Perf, device = "png", path = folder, 
       width = 16, height = 11, unit = "cm", dpi = 300)

ggsave(filename = "Rs_Perf.png", 
       plot = Rs_Perf, device = "png", path = folder, 
       width = 16, height = 11, unit = "cm", dpi = 300)

ggsave(filename = "Rc_Perf_material.png", 
       plot = Rc_Perf_material, device = "png", path = folder, 
       width = 20, height = 11, unit = "cm", dpi = 300)

ggsave(filename = "Lineplots.png", 
       plot = lineplots, device = "png", path = folder, 
       width = 17, height = 15, unit = "cm", dpi = 300)

ggsave(filename = "Barplots.png", 
       plot = barplots, device = "png", path = folder, 
       width = 17, height = 23, unit = "cm", dpi = 300)


#For Paper
Supp_fig1 <- ggarrange(lineplots, Rc_Perf_material,
                       nrow = 2, heights = c(1, 0.85),
                       labels = c("a", "b"))

ggsave(filename = "Supplementary Figure 1.png", 
       plot = Supp_fig1, device = "png", path = folder, 
       width = 17, height = 18.5, unit = "cm", dpi = 300)

ggsave(filename = "Fig1b.png", 
       plot = barplots, device = "png", path = folder, 
       width = 18, height = 16, unit = "cm", dpi = 300)
```
\newpage 
# CITATIONS 

Citations of R Studio and packages used. 

```{r}

print(citation(), style = "text")

for(i in c("rmarkdown", requiredPackages)) {
  print(i); print(citation(i), style = "text"); cat('\n')
}
```

